California Road Construction Safety Analysis: A Spatial Econometric Approach

### 2025-03-26 final/final_MoritzPeistNoemiLucchiSimonVellin.Rmd
Finalized submission

---
title: "California Road Construction Safety Analysis: A Spatial Econometric Approach"
author:
- Moritz Peist
- Noemi Lucchi
- Simon Vellin
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
    code_folding: show
    fig_caption: true
    number_sections: true
    progress: false
    warning: false
---

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE, 
  message = FALSE,
  fig.width = 8,
  fig.height = 5,
  out.width = "100%",
  fig.align = "center",
  comment = "#>"
)

@@

## 2.1 ["Effects of truck traffic on crash injury severity on rural highways in Wyoming using Bayesian binary logit models" by Ahmed et al. (2018)](https://www.sciencedirect.com/science/article/pii/S0001457518301507)

@@

## 2.2 ["Comparison of characteristics between fatal and injury accidents in the highway construction zones" by Li and Bai (2008)]( https://www.sciencedirect.com/science/article/pii/S0925753507001014)

@@

```{r data_loading}
# Load Caltrans State Highway Network
aadt <- st_read("data/Traffic_Volumes_AADT/Traffic_Volumes_AADT.shp")

@@

# Load California road network for mapping
ca_roads <- tigris::primary_roads("CA", year = 2021, refresh=TRUE) %>%
  st_transform(4326)
```

@@

@@

The temporal distribution reveals distinct patterns for accidents and construction projects. Accident rates peak in winter months (January and December) and fall (October), while construction activity is highest in spring (March-May). This inverse relationship suggests seasonal factors may influence both construction planning and accident occurrence.

@@

@@

The heatmaps confirm the spatial concentration of both accidents and construction activities in California's major urban areas. The similar patterns suggest a potential relationship, but we need to control for traffic volume differences and establish causality.

@@

Our analysis of the urban-rural breakdown is critical for understanding the heterogeneous effects of construction on accident rates.

@@

The traffic volume distribution shows a pattern likely reflecting the differences between urban and rural areas. Urban areas typically have higher traffic volumes, which could influence both construction planning and accident rates.

@@

When examining accidents by their timing relative to construction, this distribution provides initial evidence that construction activities may impact accident patterns, but we need more rigorous analysis to establish causality.

@@

The visualization illustrates the spatial relationship between construction buffers (red areas) and accident locations (blue points). Many accidents occur within or near construction zones, suggesting a potential spatial correlation.

@@

@@

Contrary to our initial hypothesis, we found significant heterogeneous effects between urban and rural environments. In urban areas, construction is associated with a more pronounced increase in accident rates (interaction coefficient = 0.163, p < 0.01), while rural areas show a potentially different pattern, suggesting that traffic management challenges differ substantially by environment.

@@

@@

# Create spatial weights - might need to adjust distance for your specific data
sp_nb <- spdep::dnearneigh(spw_coords, d1 = 0, d2 = 5000)
spw <- spdep::nb2listw(sp_nb, style = "W", zero.policy = TRUE)

@@

@@

Our spatial spillover analysis reveals nuanced effects extending beyond the immediate construction zone. Notably, areas 500m to 2km from construction sites demonstrate statistically significant increases in accident rates (coefficients ranging from 0.308 to 0.312, p < 0.001), suggesting that safety management strategies should extend well beyond the immediate construction boundaries.

@@

@@

The summary statistics reveal important patterns in accident occurrence by construction status and urban/rural setting, providing context for our main regression analyses.

@@

The models show how active construction is associated with changes in accident counts, with notable differences between urban and rural areas. The urban heterogeneity model in particular reveals important differential impacts. When examining normalized accident rates (per 10,000 vehicles), the effects may change in magnitude or significance, suggesting traffic volume plays a complex role in the relationship.

@@

The spatial model confirms the effects of construction on accident counts even after controlling for spatial dependence, providing additional evidence for the relationship.

@@

The spillover analysis reveals a pattern of effects at different distances from construction sites, with particularly notable impacts at specific distances. This suggests that construction effects aren't limited to the immediate construction zone.

@@

The event study results show patterns of accident rates before and after construction begins, providing insights into temporal dynamics and supporting or challenging the parallel trends assumption required for causal interpretation of our DiD results.

@@

The severity analysis reveals a consistent distribution of accident severity across construction periods. Approximately 87% of accidents are classified as moderate severity, with about 12% categorized as serious. Interestingly, the proportional stability suggests that while construction impacts accident frequency, it does not substantially alter the underlying severity distribution.

@@

1. **Overall Construction Effect**: Our base difference-in-differences model reveals a statistically significant increase in accident counts during construction periods (coefficient = 0.101, p < 0.001), indicating that road construction activities are associated with approximately 10.1% more accidents compared to non-construction periods..

2. **Urban-Rural Heterogeneity**: The impact of construction on accidents differs substantially between urban and rural environments. In urban areas, construction is generally associated with increased accident counts, while in rural areas, there is evidence of potentially reduced accidents during construction. This heterogeneity may be explained by several factors:
   - Higher traffic density in urban areas creates more complex interactions around construction zones
   - Rural construction may induce more cautious driving or effective traffic diversion
   - Urban construction often involves more complex traffic management challenges

3. **Spatial Spillover Effects**: The spatial spillover analysis shows that construction impacts extend beyond the immediate construction zone, with effects varying by distance. This finding has important implications for traffic management strategies, suggesting that safety measures should extend beyond the construction zone itself to areas where merging and speed adjustment typically occur.

4. **Temporal Dynamics**: The event study results provide insights into accident patterns before, during, and after construction. We observe changes in accident rates during different phases of construction, which may reflect adaptation effects as drivers adjust to changed road conditions.

5. **Severity Patterns**: Our severity analysis shows the distribution of accident severity levels (Slight, Moderate, Serious, Fatal) across different construction periods. This helps us understand whether construction not only affects accident frequency but also severity levels.

@@

1. **Targeted Resource Allocation**: Traffic management resources should be allocated differentially between urban and rural construction zones, with greater emphasis on urban areas where construction effects on accidents are typically larger.

2. **Extended Safety Zones**: Traffic safety measures should not be limited to the immediate construction zone but should extend to areas surrounding construction sites where our spillover analysis indicates significant effects.

3. **Temporal Planning**: The temporal patterns revealed in our event study suggest that enhanced safety measures may be particularly important during specific phases of construction projects.

4. **Severity-Focused Strategies**: Our severity analysis provides insights for emergency response planning and the design of safety measures aimed at mitigating more severe accidents.

@@

6. **AADT data**: The AADT data used for traffic volume normalization may include estimates or projections that do not precisely match the 2021 study period, potentially introducing minor measurement uncertainties."

>>>>

---
title: "California Road Construction Safety Analysis: A Spatial Econometric Approach"
author:
- Moritz Peist
- Noemi Lucchi
- Simon Vellin
date: "`r withr::with_locale(c('LC_TIME' = 'en_US.UTF-8'), format(Sys.Date(), '%B %d, %Y'))`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
    code_folding: show
    fig_caption: true
    number_sections: true
    progress: false
    warning: false
---

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE, 
  message = FALSE,
  progress = FALSE,
  fig.width = 8,
  fig.height = 5,
  out.width = "100%",
  fig.align = "center",
  comment = "#>"
)

@@

## ["Effects of truck traffic on crash injury severity on rural highways in Wyoming using Bayesian binary logit models" by Ahmed et al. (2018)](https://www.sciencedirect.com/science/article/pii/S0001457518301507)

@@

## ["Comparison of characteristics between fatal and injury accidents in the highway construction zones" by Li and Bai (2008)]( https://www.sciencedirect.com/science/article/pii/S0925753507001014)

@@

```{r data_loading}
# Redirect all output to a temporary file (suppresses progress bars)
sink(tempfile())

# Load Caltrans State Highway Network
aadt <- st_read("data/Traffic_Volumes_AADT/Traffic_Volumes_AADT.shp")

@@

# Load California road network for mapping
ca_roads <- tigris::primary_roads("CA", year = 2021, refresh=TRUE) %>%
  st_transform(4326)

# Restore normal output
sink()
```

@@

### Traffic Volumes
```{r AADT visualization}
# Data Preparation: Aggregate Traffic Volume
traffic_data_aggregated <- aadt %>%
  mutate(across(c(BACK_AADT, AHEAD_AADT, BACK_PEAK_, AHEAD_PEAK), 
                ~as.numeric(as.character(.)))) %>%
  # Average traffic volumes from both directions
  mutate(
    aadt = rowMeans(cbind(AHEAD_AADT, BACK_AADT), na.rm = TRUE),
    avg_peak_traffic = rowMeans(cbind(BACK_PEAK_, AHEAD_PEAK), na.rm = TRUE)
  ) %>%
  # Remove initial columns 
  select(-c(AHEAD_AADT, BACK_AADT, AHEAD_PEAK, BACK_PEAK_, AHEAD_PE_1, BACK_PEA_1))

# Visualization of Annual Average Daily Traffic
traffic_roads_plot <- ggplot() +
  # Add California road network as background
  geom_sf(data = ca_roads, color = "gray50", size = 1.5, linetype = "solid") +

  # Add traffic volume data points
  geom_sf(data = traffic_data_aggregated %>% filter(!is.na(aadt)),
          aes(color = aadt), 
          size = 2) +

  # Use viridis color palette for continuous traffic volume
  scale_color_viridis_c(
    name = "Annual Average\nDaily Traffic",
    option = "plasma",
    labels = scales::comma,
    direction = 1
  ) +

  # Zoom in to California boundaries
  coord_sf(
    xlim = c(-124.5, -114.5),  # Longitude limits for California
    ylim = c(32.5, 42.5),      # Latitude limits for California
    expand = FALSE
  ) +

  # Customize plot theme
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    legend.position = "right",
    legend.title = element_text(face = "bold")
  ) +

  # Add informative title and subtitle
  labs(
    title = "California Road Traffic Volumes",
    subtitle = "Spatial Distribution of Annual Average Daily Traffic for Post Miles",
    caption = "Data source: Caltrans State Highway Network"
  )

# Display the plot
print(traffic_roads_plot)
```
The visualization reveals the spatial distribution of Annual Average Daily Traffic (AADT) across California's road network. The highest traffic volumes (indicated by yellow and red points) are concentrated in major urban centers, particularly around Los Angeles, San Francisco, and San Diego metropolitan areas, with volumes exceeding 200,000 vehicles per day in some locations. Rural and mountainous regions show significantly lower traffic volumes (dark purple), often below 50,000 vehicles per day. This spatial variation in traffic exposure is essential for our analysis, as it provides context for normalizing accident rates and understanding how construction impacts might differ between high and low-traffic areas.

@@

The temporal distribution reveals a striking inverse relationship between accidents and construction projects throughout 2021. Accident rates show clear seasonality with peaks in January (~38,000 accidents) and December (~36,000 accidents), and a notable drop in March-April (~20,000 accidents). In contrast, construction activity peaks in March (~5,000 new projects) – precisely when accident rates are lowest – and maintains a relatively high level through June, before gradually declining toward the end of the year. This pattern suggests that seasonal factors such as weather conditions and daylight hours may influence both construction planning and accident occurrence, with transportation agencies potentially scheduling more construction during periods of historically lower accident rates.

@@

The spatial distribution maps reveal strong geographic patterns for both accidents and construction activities in California. Accidents (top map) are heavily concentrated along major transportation corridors and urban centers, with particularly high densities in the Los Angeles Basin, San Francisco Bay Area, and the Central Valley along Interstate 5. The temporal coloring shows some month-to-month variation but suggests relatively consistent spatial patterns throughout the year.

Similarly, construction sites (bottom map) follow the state's major transportation infrastructure, with higher concentrations in metropolitan areas. The severity classification shows a mix of construction types across regions, with higher severity projects (darker colors) somewhat more prevalent in urban cores. This spatial correlation between accidents and construction sites highlights the importance of controlling for location-specific factors in our causal analysis, as both are influenced by underlying traffic patterns and infrastructure density.

@@

The density heatmaps provide a striking visualization of the spatial concentration patterns for both accidents and construction activities across California. The accident heatmap (top) reveals intense clustering in major urban centers, with the Los Angeles metropolitan area showing the highest density (red), followed by the San Francisco Bay Area and San Diego (yellow-green). Several medium-intensity clusters (blue-cyan) appear along major highways in the Central Valley.

Similarly, the construction density heatmap (bottom) shows pronounced concentration in the same urban areas, with particularly high intensity (red-orange) in Los Angeles and the Bay Area. These parallel patterns strongly suggest that both accidents and construction activities follow underlying population and traffic volume distributions. This spatial correlation underscores the importance of our difference-in-differences approach with location fixed effects, which will help isolate the causal impact of construction on accidents by controlling for these shared geographic determinants.

@@

The urban-rural classification shows that approximately 66.1% of construction sites in our dataset are located in urban areas, while 33.9% are in rural areas. This distribution is essential for our heterogeneity analysis, as it ensures we have sufficient observations in both urban and rural categories to detect differential effects. The substantial representation of both categories allows us to investigate how construction impacts accident rates differently across these distinct environments, which has important implications for targeted safety measures and resource allocation.

@@

The traffic volume distribution shows substantial variation across construction sites, with a median AADT of 72,000 vehicles per day but ranging from as low as 107 to over 379,000 vehicles per day. The right-skewed distribution (evident in the first histogram) indicates that most construction sites experience moderate traffic volumes, while a smaller number of sites have extremely high traffic. The log-scale visualization reveals a more nuanced bimodal pattern, suggesting two distinct categories of roads: lower-volume rural or suburban routes (clustered around 10^3-10^4 vehicles/day) and higher-volume urban arterials and highways (clustered around 10^5 vehicles/day). This considerable variation in traffic exposure is critical to control for in our analysis, as it directly affects the accident risk baseline across different locations.

@@

The distribution of accidents by construction period reveals a notable pattern: 46.2% of accidents within our buffer zones occurred before construction began, 24.1% occurred during active construction, and 29.7% occurred after construction ended. This preliminary finding suggests potential changes in accident patterns associated with construction activities. The relatively lower percentage during construction could indicate several possibilities: (1) driver adaptation to construction zones through increased caution, (2) reduced traffic volume due to drivers avoiding construction areas, or (3) effective traffic management measures during construction. However, this descriptive analysis doesn't control for other factors like location characteristics or seasonal patterns, highlighting the need for our more rigorous difference-in-differences approach.

@@

The visualization of the Los Angeles area provides a clear illustration of the spatial relationship between construction buffers (red circles) and accident locations (blue points). Many accidents cluster within or near the 500-meter construction buffer zones, particularly along major transportation corridors. We observe that some buffers contain dense accident clusters, while others have relatively few accidents, suggesting potential heterogeneity in the impact of construction on accident frequency. This visualization highlights the importance of our buffer-based analytical approach, which allows us to capture localized effects of construction activities on nearby accident patterns while controlling for location-specific factors through fixed effects.

@@

The base difference-in-differences model provides compelling evidence for the causal impact of construction on accident rates. The coefficient for active construction (treatment) is 0.101 (p < 0.001), indicating that during construction periods, the average number of accidents in buffer zones increases by approximately 0.101 accidents per month compared to non-construction periods. With clustered standard errors at the buffer level, this effect is highly statistically significant, suggesting a robust relationship. The low within-R² (9.001e-5) is common in fixed effects models with binary treatment variables and indicates that while construction activity explains a small portion of the overall variation in accidents, its effect is statistically meaningful and precisely estimated.

@@

The heterogeneity analysis reveals striking differences in how construction affects accidents across urban and rural environments. The base treatment effect coefficient (-0.005) is not statistically significant, suggesting minimal impact of construction on accidents in rural areas. However, the interaction term (treatment × is_urbanTRUE) shows a large and highly significant coefficient of 0.163 (p < 0.01), indicating that in urban areas, construction is associated with an additional 0.163 accidents per month beyond any rural effect. This substantial urban-rural disparity (with urban areas experiencing approximately 16.3% more accidents during construction) suggests that traffic management challenges differ fundamentally between these environments, with urban construction zones posing significantly greater accident risks that require targeted mitigation strategies.

@@

When examining normalized accident rates per 10,000 vehicles, we observe different patterns than in the absolute count models. In the base rate model, the treatment effect is positive but not statistically significant (0.013, p=0.41), suggesting that after controlling for traffic exposure, construction's impact on accident rates is more nuanced.

However, the urban heterogeneity model for normalized rates shows a similar pattern to our count model: the base treatment effect for rural areas is negative (-0.059, p=0.12) though not statistically significant, while the urban interaction term is positive and significant (0.110, p<0.01). This indicates that in urban areas, construction is associated with an increase of 0.110 accidents per 10,000 vehicles compared to rural areas. These normalized models provide important context by accounting for traffic volume differences, suggesting that the urban-rural disparity in construction impacts persists even after controlling for exposure levels.

@@

# Create spatial weights - might need to adjust distance for our specific data
sp_nb <- spdep::dnearneigh(spw_coords, d1 = 0, d2 = 5000)
spw <- spdep::nb2listw(sp_nb, style = "W", zero.policy = TRUE)

@@

The spatial lag model confirms the presence of significant spatial autocorrelation in accident patterns, with a spatial lag parameter (ρ) of 0.765 (p < 0.001). This high and statistically significant value indicates strong spatial clustering in accident occurrences, where accident rates in one location are substantially influenced by rates in neighboring areas.

After accounting for this spatial dependence, the treatment effect of construction remains positive and significant (0.622, p < 0.001), providing additional evidence for our main finding. The model also identifies a negative effect of urban classification (-0.256, p < 0.01), likely reflecting the fixed characteristics of urban areas that are captured in this cross-sectional spatial model rather than the panel approach. The significant positive coefficient for traffic volume (0.438, p < 0.001) confirms the expected relationship between traffic exposure and accident frequency. This spatial analysis strengthens our causal claims by demonstrating that construction effects persist even after controlling for complex spatial dependencies.

@@

Our spatial spillover analysis reveals nuanced effects extending beyond the immediate construction zone. Using the ">5km" distance band as reference, we find that areas within 0-500m of construction sites show a negative but non-significant effect (-0.064, p=0.93), which may reflect successful traffic management or diversion within immediate construction boundaries.

Notably, the strongest positive effects appear in the 500-1000m band (0.308, p<0.001) and 1-2km band (0.312, p<0.001), indicating significantly higher accident rates in these transition zones where traffic flow and driver behavior adjust to approaching or leaving construction. Even in the 2-5km band, we observe a smaller but still significant effect (0.139, p<0.001). These findings suggest that safety management strategies should extend well beyond immediate construction boundaries, with particular attention to transition zones where accident risk may actually be higher than at the construction site itself due to changing traffic patterns and driver adaptation.

@@

The event study results provide temporal context for our difference-in-differences estimates. Using the period 3 months before construction ("-3to0m") as the reference, we observe that locations destined for construction show significantly lower accident rates far before construction begins ("<-12m" coefficient = -0.690, p<0.001), with this difference gradually diminishing as construction approaches.

During and after construction (0to3m, 3to6m, 6to12m, >12m), coefficients are close to zero and not statistically significant, suggesting accident rates similar to the pre-construction reference period. This pattern supports the parallel trends assumption required for causal interpretation of our DiD results, as we don't see diverging pre-trends immediately before treatment. The significant difference in very early periods may reflect long-term planning effects, where areas selected for eventual construction have different characteristics. Overall, these results strengthen our causal claims by demonstrating appropriate temporal patterns in the data.

@@

The summary statistics reveal stark differences in accident patterns across urban-rural divides and construction status. In rural areas without construction (FALSE, FALSE), the mean accident count is 0.647 with a median of 0, while in urban areas without construction (FALSE, TRUE), the mean jumps to 4.383 with a median of 1 – nearly seven times higher. Similarly, during active construction, urban areas (TRUE, TRUE) show mean accident counts of 4.672 compared to 0.783 in rural areas (TRUE, FALSE).

Notably, when examining normalized accident rates per 10,000 vehicles, the urban-rural gap narrows substantially: rural areas show rates of 0.986 (without construction) and 0.827 (with construction), while urban areas show rates of 0.395 (without construction) and 0.448 (with construction). This suggests that higher urban accident counts are partly explained by higher traffic volumes. The data also reveals that construction appears to reduce normalized accident rates in rural areas (-16%) while increasing them in urban areas (+13%), providing preliminary evidence for heterogeneous effects that will be further examined in our regression models.

@@

The difference-in-differences models provide robust evidence for the causal impact of construction on accident rates, with important heterogeneity across urban and rural environments. The base model shows a significant positive effect of construction (treatment = 0.101, p < 0.001), indicating that construction activities increase accident counts by approximately 10.1% on average.

When examining urban-rural heterogeneity, we find that the effect is concentrated in urban areas: while the base treatment effect (representing rural areas) is actually negative but insignificant (-0.005), the urban interaction term is strongly positive (0.163, p < 0.01), indicating substantially higher accident rates during construction in urban environments.

The normalized rate models (controlling for traffic volume) show complementary results, with the urban interaction term remaining significant (0.110, p < 0.01) even after traffic exposure normalization. This strengthens our conclusion that urban and rural areas experience fundamentally different safety impacts from construction activities, with urban construction zones posing greater accident risks that cannot be explained solely by traffic volume differences.

@@

The spatial lag model provides strong evidence for both spatial dependencies in accident patterns and the impact of construction activities. The spatial lag parameter (ρ = 0.765, p < 0.001) confirms significant spatial autocorrelation, indicating that accident rates in one location are strongly influenced by neighboring areas. Even after accounting for this spatial clustering, the treatment effect of construction remains positive and highly significant (0.622, p < 0.001), providing additional support for our main finding.

Interestingly, in this cross-sectional spatial model, the Urban coefficient is negative (-0.256, p = 0.001), likely capturing fixed characteristics of urban areas rather than the treatment interaction effects seen in our panel models. The significant positive coefficient for traffic volume (0.438, p < 0.001) confirms the expected relationship between traffic exposure and accident frequency. With a log-likelihood of -101588.9 and AIC of 203190.0, the model demonstrates good fit, strengthening our confidence in the identified spatial relationships and treatment effects.

@@

The spillover analysis reveals a non-intuitive spatial pattern of construction effects on accident rates. Using areas beyond 5km as the reference category, we observe that the immediate construction zone (0-500m) shows a slight negative but non-significant effect (-0.064, p=0.93), possibly reflecting successful traffic management or driver avoidance of these areas.

However, the strongest positive effects appear in transition zones: the 500-1000m band (0.308, p<0.001) and 1-2km band (0.312, p<0.001) show statistically significant increases in accident counts, with magnitudes over three times larger than our base model estimate. Even in the 2-5km band, we observe a smaller but still significant effect (0.139, p<0.001).

This spatial pattern suggests that accident risk may actually peak in areas approaching or leaving construction zones rather than within them, where speed changes, lane merges, and driver adjustment to changing conditions create hazardous transition environments. These findings have important implications for traffic management, indicating that safety measures should extend well beyond construction boundaries with particular attention to these high-risk transition zones.

@@

The event study results reveal important temporal dynamics in accident patterns around construction periods. Using the 3 months before construction start as the reference period, we observe that areas selected for construction have significantly lower accident rates long before construction begins (coefficient for "<-12m" = -0.690, p<0.001), with this difference gradually diminishing as construction approaches.

During and after construction (periods "0to3m" through ">12m"), coefficients remain close to zero and statistically insignificant, indicating accident rates similar to the immediate pre-construction period. This pattern supports two important conclusions: (1) the parallel trends assumption necessary for causal interpretation of our DiD estimates is reasonably satisfied in the periods immediately before treatment, and (2) the long-term pre-construction difference suggests selection effects in areas chosen for construction projects.

These temporal dynamics provide context for interpreting our main results, suggesting that while construction causes a relative increase in accidents compared to the immediate pre-construction period, accident rates might still be lower than in typical non-construction areas due to selection effects. This nuanced interpretation strengthens our analysis by accounting for temporal heterogeneity in treatment effects.

@@

The severity analysis reveals a remarkably consistent distribution of accident severity levels across different construction periods. Approximately 87% of accidents are classified as moderate severity (level 2) across all periods, with around 12% categorized as serious (level 3) and less than 1% as either slight (level 1) or fatal (level 4).

This stability suggests that while construction impacts accident frequency, as demonstrated in our main models, it does not substantially alter the proportional severity distribution. This finding has important implications for safety planning, indicating that construction zones don't necessarily produce more severe accidents but rather increase the overall number of accidents while maintaining similar severity patterns. Traffic management strategies should therefore focus primarily on reducing accident occurrence rather than specifically targeting severity reduction, as the underlying severity distribution appears to be a stable characteristic of the road network and driver behavior.

@@

1. **Overall Construction Effect**: Our base difference-in-differences model reveals a statistically significant increase in accident counts during construction periods (coefficient = 0.101, p < 0.001), indicating that road construction activities are associated with approximately 10.1% more accidents compared to non-construction periods.

2. **Urban-Rural Heterogeneity**: The impact of construction on accidents differs substantially between urban and rural environments. In urban areas, construction is associated with a significant increase in accident counts (interaction coefficient = 0.163, p < 0.01), while in rural areas, there is evidence of potentially reduced accidents during construction (base coefficient = -0.005, though not statistically significant). This heterogeneity may be explained by several factors:
   - Higher traffic density in urban areas creates more complex interactions around construction zones
   - Rural construction may induce more cautious driving or effective traffic diversion
   - Urban construction often involves more complex traffic management challenges

3. **Spatial Spillover Effects**: The spatial spillover analysis shows that construction impacts extend beyond the immediate construction zone, with effects varying by distance. Areas 500m-2km from construction sites show the largest increases in accident rates (coefficients 0.308-0.312, p < 0.001), while immediate construction zones (0-500m) show minimal effects. This suggests that transition zones where traffic adjusts to approaching or leaving construction are particularly high-risk areas that warrant targeted safety measures.

4. **Temporal Dynamics**: The event study results show that locations selected for construction have different accident patterns long before construction begins, with accident rates gradually approaching the reference period as construction nears. This suggests potential selection effects in areas chosen for construction projects and provides context for interpreting our main treatment effects.

5. **Severity Patterns**: Our severity analysis shows a consistent distribution of accident severity levels across different construction periods (approximately 87% moderate, 12% serious), indicating that construction affects accident frequency but not the underlying severity distribution.

@@

1. **Targeted Resource Allocation**: Traffic management resources should be allocated differentially between urban and rural construction zones, with greater emphasis on urban areas where construction effects on accidents are typically larger (coefficient = 0.163, p < 0.01 for urban interaction). Rural construction zones may require less intensive intervention given their potentially neutral or even positive safety effects.

2. **Extended Safety Zones**: Traffic safety measures should not be limited to the immediate construction zone but should extend to areas surrounding construction sites, particularly the 500m-2km transition zones where our spillover analysis indicates significant effects (coefficients 0.308-0.312, p < 0.001). This could include enhanced signage, speed management, and temporary traffic calming measures in these approach and departure areas.

3. **Temporal Planning**: The temporal patterns revealed in our event study suggest that enhanced safety measures may be particularly important during specific phases of construction projects. The relatively stable accident rates during and after construction compared to the immediate pre-construction period suggest that current construction management practices may be partially effective but could be further improved.

4. **Severity-Focused Strategies**: Our severity analysis provides insights for emergency response planning and the design of safety measures. Since the severity distribution remains consistent across construction periods, general accident prevention measures rather than severity-specific interventions may be most effective.

@@

6. **AADT data**: The AADT data used for traffic volume normalization may include estimates or projections that do not precisely match the 2021 study period, potentially introducing minor measurement uncertainties.
